---
title: "DOE Student Listings"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(magrittr)
R.utils::sourceDirectory("~/git/vecf_counts/src/functions/")
source(file = "~/git/vecf_counts/src/functions/fix_column_names.R")
source(file = "~/git/vecf_counts/src/functions/is_blank.R")
library(dplyr)

cust_by_yr <- fix_column_names(data.table::fread("~/git/vecf_counts/data/original/q4/DOE/Unique Students Listing.csv", colClasses = "character"))
values <- fix_column_names(data.table::fread("~/git/vecf_counts/data/original/ExportedValues.csv"))
colnames(cust_by_yr)
```

####Notes: 

* I skipped [match_type] and [associated_id].  
+ match_type - missing in metadata, 100% of records have value "D"
+ associated_id - assuming this is an internal id? 
* five observations had school numbers that didn't appear in the data dictionary 

```{r columns, echo=FALSE}
#. unique_id ----
pct_complete_unique_id <- (nrow(cust_by_yr) - cust_by_yr[is_blank(unique_id), .N]) / nrow(cust_by_yr)
#.. unique value count ----
unq_count_unique_id <- nrow(unique(cust_by_yr[, .(unique_id)]))
#.. valid values ----
wrong_id_length <- cust_by_yr[nchar(unique_id) != 7, .N]
pct_valid_values_unique_id <- (nrow(cust_by_yr) - wrong_id_length) / nrow(cust_by_yr)

#. birth month & year ----
pct_complete_birth_m <- (nrow(cust_by_yr) - cust_by_yr[is_blank(birth_month), .N]) / nrow(cust_by_yr)
pct_complete_birth_y <- (nrow(cust_by_yr) - cust_by_yr[is_blank(birth_year), .N]) / nrow(cust_by_yr)
#.. unique value count ----
unq_count_birth_m <- nrow(unique(cust_by_yr[, .(birth_month)]))
unq_count_birth_y <- nrow(unique(cust_by_yr[, .(birth_year)]))
#.. valid values ----
pct_valid_values_birth_m <- (nrow(cust_by_yr) - cust_by_yr[as.numeric(birth_month) < 1| as.numeric(birth_month) > 12, .N]) / nrow(cust_by_yr)
pct_valid_values_birth_y <- (nrow(cust_by_yr) - cust_by_yr[as.numeric(birth_year) < 2010, .N]) / nrow(cust_by_yr)

#. demographics columns - gender, race, english learner ----
pct_complete_gend <- (nrow(cust_by_yr) - cust_by_yr[is_blank(gender), .N]) / nrow(cust_by_yr)
pct_complete_race <- (nrow(cust_by_yr) - cust_by_yr[is_blank(race_type), .N]) / nrow(cust_by_yr)
pct_complete_englearn <- (nrow(cust_by_yr) - cust_by_yr[is_blank(funded_english_learner_flag), .N]) / nrow(cust_by_yr)
#.. unique value count ----
unq_count_gend <- nrow(unique(cust_by_yr[, .(gender)]))
unq_count_race <- nrow(unique(cust_by_yr[, .(race_type)]))
unq_count_englearn <- nrow(unique(cust_by_yr[, .(funded_english_learner_flag)]))
#.. valid values ----
pct_valid_values_gend <- (nrow(cust_by_yr) - cust_by_yr[!gender %in% c("F", "M"), .N]) / nrow(cust_by_yr)
pct_valid_values_race <- (nrow(cust_by_yr) - cust_by_yr[as.numeric(race_type) == 0, .N]) / nrow(cust_by_yr)
pct_valid_values_englearn <- (nrow(cust_by_yr) - cust_by_yr[!funded_english_learner_flag %in% c("N", "Y"), .N]) / nrow(cust_by_yr)

#. school columns - school year, entry date, grade code, school division ----
pct_complete_sch_yr <- (nrow(cust_by_yr) - cust_by_yr[is_blank(school_year), .N]) / nrow(cust_by_yr)
pct_complete_entry <- (nrow(cust_by_yr) - cust_by_yr[is_blank(entry_date), .N]) / nrow(cust_by_yr)
pct_complete_grade <- (nrow(cust_by_yr) - cust_by_yr[is_blank(grade_code), .N]) / nrow(cust_by_yr)
pct_complete_sch_num <- (nrow(cust_by_yr) - cust_by_yr[is_blank(division_number_reporting_school_numbe), .N]) / nrow(cust_by_yr)
#.. unique value count ----
unq_count_sch_yr <- nrow(unique(cust_by_yr[, .(school_year)]))
unq_count_entry <- nrow(unique(cust_by_yr[, .(entry_date)]))
unq_count_grade <- nrow(unique(cust_by_yr[, .(grade_code)]))
unq_count_sch_num <- nrow(unique(cust_by_yr[, .(division_number_reporting_school_numbe)]))
#.. valid values ----
pct_valid_values_sch_yr <- (nrow(cust_by_yr) - cust_by_yr[as.numeric(school_year) < 2013, .N]) / nrow(cust_by_yr)
pct_valid_values_entry <- (nrow(cust_by_yr) - cust_by_yr[as.numeric(stringr::str_sub(entry_date, start = 1L, end = 4L)) <2013, .N]) / nrow(cust_by_yr)
pct_valid_values_grade <- (nrow(cust_by_yr) - cust_by_yr[grade_code %in% c("KG", "PK"), .N]) / nrow(cust_by_yr)
valid_sch_num <- values[agency_name == "Virginia Department of Education"][view_name == "Student Records"][column_name == "Division Number - Reporting School Number"][,value]
pct_valid_values_sch_num <- (nrow(cust_by_yr) - cust_by_yr[!division_number_reporting_school_numbe %in% valid_sch_num, .N]) / nrow(cust_by_yr)


```


```{r UniqueID, echo = FALSE}

SummaryEval <- 
  rbind(
    tibble::tibble("Column" = "ID", "Measure" = c("% Complete", "# Unique", "% Valid "), 
                         "Value" = c(pct_complete_unique_id, unq_count_unique_id, pct_valid_values_unique_id)),
    tibble::tibble("Column" = "Birth Month", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_birth_m, unq_count_birth_m, pct_valid_values_birth_m)),
    tibble::tibble("Column" = "Birth Year", "Measure"=c("% Complete", "# Unique", "% Valid "),
               "Value" = c(pct_complete_birth_y,unq_count_birth_y, pct_valid_values_birth_y)),
    tibble::tibble("Column" = "School Yr", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_sch_yr, unq_count_sch_yr, pct_valid_values_sch_yr)),
    
    tibble::tibble("Column" = "Entry Date", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_entry, unq_count_entry, pct_valid_values_entry)),
    tibble::tibble("Column" = "English", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_englearn, unq_count_englearn, pct_valid_values_englearn)),
    tibble::tibble("Column" = "Gender", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_gend, unq_count_gend, pct_valid_values_gend)),
    tibble::tibble("Column" = "Grade", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_grade, unq_count_grade, pct_valid_values_grade)),
    tibble::tibble("Column" = "Race Type", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_race, unq_count_race, pct_valid_values_race)),
    tibble::tibble("Column" = "School No", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_sch_num, unq_count_sch_num, pct_valid_values_sch_num))
    
    ) #rbind  

```

### Summary Table

```{R, echo = FALSE}
knitr::kable(reshape2::recast(SummaryEval, Column~Measure))
```

####Validations: 
* unique_id - filtered for nchar = 7
* birth_month -filtered for months betweeen 1 and 12 
* birth_year - filtered for years after 2010
* school_year - filtered for years after 2013 
* entry_date - filtered for years after 2013
* funded_english_learner_flag - filtered for values other than "N" and "Y"
* gender - filtered for values other than "M" and "F"
* grade_code - filtered for values other than "PK" (prekindergarten) and "KG" (kindergarten)
* race_type - filtered out value "0" indicating race unknown (note value 99 representing multiple race came into use in 2010-2011)
* division_number_reporting_school_numbe - filtered for codes appearing in the dictionary

####Longitudinal checks

Total number of records is:

```{r, echo = FALSE}
print(nrow(cust_by_yr))
```

Since there are 135 unique IDs - no need to check persistence of traits for the duplicates? 


