---
title: "DSS - Customers"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(magrittr)
R.utils::sourceDirectory("~/git/vecf_counts/src/functions/")
source(file = "~/git/vecf_counts/src/functions/fix_column_names.R")
source(file = "~/git/vecf_counts/src/functions/is_blank.R")

cust_by_yr <- fix_column_names(data.table::fread("~/git/vecf_counts/data/original/q4/DSS/DSS Customers By Year.csv", colClasses = "character"))
values <- fix_column_names(data.table::fread("~/git/vecf_counts/data/original/ExportedValues.csv"))

```

####Notes: 

* I skipped [study_group_id]

```{r columns, echo=FALSE}
#. unique_id ----
pct_complete_unique_id <- (nrow(cust_by_yr) - cust_by_yr[is_blank(unique_id), .N]) / nrow(cust_by_yr)
#.. unique value count ----
unq_count_unique_id <- nrow(unique(cust_by_yr[, .(unique_id)]))
#.. valid values ----
wrong_id_length <- cust_by_yr[nchar(unique_id) != 7, .N]
pct_valid_values_unique_id <- (nrow(cust_by_yr) - wrong_id_length) / nrow(cust_by_yr)

#. age columns ---- 
#.. completeness ----
pct_complete_age_class_code <- (nrow(cust_by_yr) - cust_by_yr[is_blank(age_class_code), .N]) / nrow(cust_by_yr)
pct_complete_age_group_code <- (nrow(cust_by_yr) - cust_by_yr[is_blank(age_group_code), .N]) / nrow(cust_by_yr)
pct_complete_age_type_code <- (nrow(cust_by_yr) - cust_by_yr[is_blank(age_type_code), .N]) / nrow(cust_by_yr)
pct_complete_month_of_birth <- (nrow(cust_by_yr) - cust_by_yr[is_blank(month_of_birth), .N]) / nrow(cust_by_yr)
pct_complete_year_of_birth <- (nrow(cust_by_yr) - cust_by_yr[is_blank(year_of_birth), .N]) / nrow(cust_by_yr)
#.. unique value count ----
unq_count_age_class_code <- nrow(unique(cust_by_yr[, .(age_class_code)]))
unq_count_age_group_code <- nrow(unique(cust_by_yr[, .(age_group_code)]))
unq_count_age_type_code <- nrow(unique(cust_by_yr[, .(age_type_code)]))
unq_count_month_of_birth <- nrow(unique(cust_by_yr[, .(month_of_birth)]))
unq_count_year_of_birth <- nrow(unique(cust_by_yr[, .(year_of_birth)]))
#.. valid values ----
pct_valid_values_age_class_code <- (nrow(cust_by_yr) - cust_by_yr[is_blank(age_class_code)|age_class_code %in% c("90", "91"), .N]) / nrow(cust_by_yr)
pct_valid_values_age_group_code <- (nrow(cust_by_yr) - cust_by_yr[is_blank(age_group_code)|age_group_code %in% c("90", "91"), .N]) / nrow(cust_by_yr)
pct_valid_values_age_type_code <- (nrow(cust_by_yr) - cust_by_yr[is_blank(age_class_code)|age_type_code %in% c("90", "91"), .N]) / nrow(cust_by_yr)
pct_valid_values_month_of_birth <- (nrow(cust_by_yr) - cust_by_yr[is_blank(month_of_birth)|!as.numeric(month_of_birth) %in% c(1:12), .N]) / nrow(cust_by_yr)
pct_valid_values_year_of_birth <- (nrow(cust_by_yr) - cust_by_yr[is_blank(year_of_birth)|!as.integer(year_of_birth) >= 1900, .N]) / nrow(cust_by_yr) 

#. school type ----
pct_complete_school_type <- (nrow(cust_by_yr) - cust_by_yr[is_blank(school_type_code), .N]) / nrow(cust_by_yr)
#.. unique value count ----
unq_count_school_type <- nrow(unique(cust_by_yr[, .(school_type_code)]))
#.. valid values ----
wrong_sch_type <- cust_by_yr[is_blank(school_type_code) | school_type_code == "0", .N]
pct_valid_values_sch_type <- (nrow(cust_by_yr) - wrong_sch_type) / nrow(cust_by_yr)

#. ethnicity ----
pct_complete_ethnicity <- (nrow(cust_by_yr) - cust_by_yr[is_blank(ethnicity_code), .N]) / nrow(cust_by_yr)
#.. unique value count ----
unq_count_ethnicity <- nrow(unique(cust_by_yr[, .(ethnicity_code)]))
#.. valid values ----
valid_eth <- values[agency_name == "Department of Social Services"][view_name == "DSS Customers By Year"][column_name == "Ethnicity Code"][,value]
pct_valid_values_ethnicity <- (nrow(cust_by_yr) - cust_by_yr[!ethnicity_code %in% valid_eth|ethnicity_code=="0", .N]) / nrow(cust_by_yr)

#. gender ----
pct_complete_gender <- (nrow(cust_by_yr) - cust_by_yr[is_blank(gender_code), .N]) / nrow(cust_by_yr)
#.. unique value count ----
unq_count_gender <- nrow(unique(cust_by_yr[, .(gender_code)]))
#.. valid values ----
invalid_gend <- c("1", "2")
pct_valid_values_gender <- (nrow(cust_by_yr) - cust_by_yr[!gender_code %in% invalid_gend, .N]) / nrow(cust_by_yr)

#. race columns ---- 
#.. completeness ----
pct_complete_race_amer_ind <- (nrow(cust_by_yr) - cust_by_yr[is_blank(cust_race_is_amer_indian_alaska_native_ind), .N]) / nrow(cust_by_yr)
pct_complete_race_asian <- (nrow(cust_by_yr) - cust_by_yr[is_blank(customer_race_is_asisan_indicator), .N]) / nrow(cust_by_yr)
pct_complete_race_black <- (nrow(cust_by_yr) - cust_by_yr[is_blank(customer_race_is_black_indicator), .N]) / nrow(cust_by_yr)
pct_complete_race_pacific <- (nrow(cust_by_yr) - cust_by_yr[is_blank(cust_race_is_hawaiian_pacific_islander_ind), .N]) / nrow(cust_by_yr)
pct_complete_race_other <- (nrow(cust_by_yr) - cust_by_yr[is_blank(customer_race_is_other_indicator), .N]) / nrow(cust_by_yr)
pct_complete_race_white <- (nrow(cust_by_yr) - cust_by_yr[is_blank(customer_race_is_white_indicator), .N]) / nrow(cust_by_yr)
#.. unique value count ----
unq_count_race_amer_ind <- nrow(unique(cust_by_yr[, .(cust_race_is_amer_indian_alaska_native_ind)]))
unq_count_race_asian <- nrow(unique(cust_by_yr[, .(customer_race_is_asisan_indicator)]))
unq_count_race_black <- nrow(unique(cust_by_yr[, .(customer_race_is_black_indicator)]))
unq_count_race_pacific <- nrow(unique(cust_by_yr[, .(cust_race_is_hawaiian_pacific_islander_ind)]))
unq_count_race_other <- nrow(unique(cust_by_yr[, .(customer_race_is_other_indicator)]))
unq_count_race_white <- nrow(unique(cust_by_yr[, .(customer_race_is_white_indicator)]))
#.. valid values ----
pct_valid_values_race_amer_ind <- (nrow(cust_by_yr) - nrow(cust_by_yr[!cust_race_is_amer_indian_alaska_native_ind %in% c("N", "Y")])) / nrow(cust_by_yr)
pct_valid_values_race_asian <- (nrow(cust_by_yr) - nrow(cust_by_yr[!customer_race_is_asisan_indicator %in% c("N", "Y")])) / nrow(cust_by_yr)
pct_valid_values_race_black <- (nrow(cust_by_yr) - nrow(cust_by_yr[!customer_race_is_black_indicator %in% c("N", "Y")]))/ nrow(cust_by_yr)
pct_valid_values_race_pacific <- (nrow(cust_by_yr) - nrow(cust_by_yr[!cust_race_is_hawaiian_pacific_islander_ind %in% c("N", "Y")])) / nrow(cust_by_yr)
pct_valid_values_race_other <- (nrow(cust_by_yr) - nrow(cust_by_yr[!customer_race_is_other_indicator%in% c("N", "Y")])) / nrow(cust_by_yr)
pct_valid_values_race_white <- (nrow(cust_by_yr) - nrow(cust_by_yr[!customer_race_is_white_indicator %in% c("N", "Y")])) / nrow(cust_by_yr)


#. foster indicator ----
pct_complete_foster <- (nrow(cust_by_yr) - cust_by_yr[is_blank(foster_care_case_indicator), .N]) / nrow(cust_by_yr)
#.. unique value count ----
unq_count_foster <- nrow(unique(cust_by_yr[, .(foster_care_case_indicator)]))
#.. valid values ----
invalid_ind <- c("N", "Y")
pct_valid_values_foster <- (nrow(cust_by_yr) - cust_by_yr[!foster_care_case_indicator %in% invalid_ind, .N]) / nrow(cust_by_yr)

#. SNAP indicator ----
pct_complete_snap <- (nrow(cust_by_yr) - cust_by_yr[is_blank(snap_case_indicator), .N]) / nrow(cust_by_yr)
#.. unique value count ----
unq_count_snap <- nrow(unique(cust_by_yr[, .(snap_case_indicator)]))
#.. valid values ----
pct_valid_values_snap <- (nrow(cust_by_yr) - cust_by_yr[!snap_case_indicator %in% invalid_ind, .N]) / nrow(cust_by_yr)

#. TANF indicator ----
pct_complete_TANF <- (nrow(cust_by_yr) - cust_by_yr[is_blank(tanf_case_indicator), .N]) / nrow(cust_by_yr)
#.. unique value count ----
unq_count_TANF <- nrow(unique(cust_by_yr[, .(tanf_case_indicator)]))
#.. valid values ----
pct_valid_values_TANF <- (nrow(cust_by_yr) - cust_by_yr[!tanf_case_indicator %in% invalid_ind, .N]) / nrow(cust_by_yr)

#. calendar year ----
pct_complete_cal <- (nrow(cust_by_yr) - cust_by_yr[is_blank(calender_year_number), .N]) / nrow(cust_by_yr)
#.. unique value count ----
unq_count_cal <- nrow(unique(cust_by_yr[, .(calender_year_number)]))
#.. valid values ----
pct_valid_values_cal <- (nrow(cust_by_yr) - cust_by_yr[!as.integer(calender_year_number) < 2013, .N]) / nrow(cust_by_yr)

```


```{r UniqueID, echo = FALSE}
SummaryEval <-
  rbind(
  tibble::tibble("Column" = "ID", "Measure" = c("% Complete", "# Unique", "% Valid "), 
                         "Value" = c(pct_complete_unique_id, 
                                     unq_count_unique_id, pct_valid_values_unique_id)),
    tibble::tibble("Column" = "Age Class", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_age_class_code, 
                               unq_count_age_class_code, pct_valid_values_age_class_code)),
    tibble::tibble("Column" = "Age Group", "Measure"=c("% Complete", "# Unique", "% Valid "),
               "Value" = c(pct_complete_age_group_code, 
                           unq_count_age_group_code, pct_valid_values_age_group_code)),
    tibble::tibble("Column" = "Age Type", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_age_group_code, 
                               unq_count_age_type_code, pct_valid_values_age_type_code)),
    tibble::tibble("Column" = "Month of Birth", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_age_group_code, 
                               unq_count_month_of_birth, pct_valid_values_month_of_birth)),
    tibble::tibble("Column" = "Year of Birth", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_year_of_birth, 
                               unq_count_year_of_birth, pct_valid_values_year_of_birth)),
    tibble::tibble("Column" = "School Type", "Measure" = c("% Complete", "# Unique", "% Valid "), 
                          "Value" = c(pct_complete_school_type, unq_count_school_type, pct_valid_values_sch_type)),
    tibble::tibble("Column" = "Ethnicity", "Measure" = c("% Complete", "# Unique", "% Valid "), 
                          "Value" = c(pct_complete_ethnicity, unq_count_ethnicity, pct_valid_values_ethnicity)),
    tibble::tibble("Column" = "Gender", "Measure" = c("% Complete", "# Unique", "% Valid "), 
                          "Value" = c(pct_complete_gender, unq_count_gender, pct_valid_values_gender)),
    tibble::tibble("Column" = "Race: Amer Ind", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_race_amer_ind, unq_count_race_amer_ind, 
                               pct_valid_values_race_amer_ind)),
    tibble::tibble("Column" = "Race: Asian", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_race_asian, unq_count_race_asian, 
                               pct_valid_values_race_asian)),
    tibble::tibble("Column" = "Race: Black", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_race_black, unq_count_race_black, 
                               pct_valid_values_race_black)),
    tibble::tibble("Column" = "Race: Pacific", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_race_pacific, unq_count_race_pacific, 
                               pct_valid_values_race_pacific)),
    tibble::tibble("Column" = "Race: Other", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_race_other, unq_count_race_other, 
                               pct_valid_values_race_other)),
    tibble::tibble("Column" = "Race: White", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_race_white, unq_count_race_white, 
                               pct_valid_values_race_white)),
    tibble::tibble("Column" = "Foster", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_foster, unq_count_foster, 
                               pct_valid_values_foster)),
    tibble::tibble("Column" = "SNAP", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_snap, unq_count_snap, 
                               pct_valid_values_snap)),
    tibble::tibble("Column" = "TANF", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_TANF, unq_count_TANF, 
                               pct_valid_values_TANF)),
    tibble::tibble("Column" = "Service Year", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_cal, unq_count_cal, 
                               pct_valid_values_cal))
  ) #rbind

```

### Summary Table

```{R}
knitr::kable(reshape2::recast(SummaryEval, Column~Measure))
```


####Validations: 
* unique_id - filtered for nchar = 7 
* age_class_code - filtered out values 90 (not valid) and 91 (source blank) 
* age_group_code - filtered out values 90 and 91
* age_type_code  - filtered out values 90 and 91
* month_of_birth - filtered out blanks and values other than 1:12
* year_of_birth  - filtered out values before 1900
* school_type_code - filtered out school type 0 (source blank)   
* ethnicity_code - filtered out values that do not appear in the dictionary (3, 90, 93) and value 0 (source blank)
* foster_care_case_indicator - filtered out values other than "N" and "Y"
* gender_code  - filtered out values other than "1" (femal) and "2" (male)
* cust_race_is_amer_indian_alaska_native_ind - filtered out values other than "N" and "Y"
* customer_race_is_asisan_indicator  - filtered out values other than "N" and "Y"
* customer_race_is_black_indicator  - filtered out values other than "N" and "Y"
* cust_race_is_hawaiian_pacific_islander_ind - filtered out values other than "N" and "Y"
* customer_race_is_other_indicator  - filtered out values other than "N" and "Y"
* customer_race_is_white_indicator  - filtered out values other than "N" and "Y"
* snap_case_indicator - filtered out values other than "N" and "Y"
* tanf_case_indicator  - filtered out values other than "N" and "Y"
* calender_year_number - filtered out values before 2013

####Longitudinal checks

Total number of records is:

```{r, echo = FALSE}
print(nrow(cust_by_yr))
```


There are 1,325,911 duplicates by unique id
There are 1,325,867 duplicates by unique id + year of birth
There are 1,325,844 duplicates by unique id + year of birth + month of birth
There are 1,325,566 duplicates by unique id + year of birth + month of birth + gender
There are 1,323,281 duplicates by unique id + year of birth + month of birth + gender + ethnicity


Using the smallest level of duplication, the percent unduplication rate is:  
```{r, echo = FALSE}
# duplicated_entries <- cust_by_yr[, .N, by = .(unique_id)][N>1, .(unique_id)] 
# 1,325,911  duplicates
# duplicated_entries <- cust_by_yr[, .N, by = .(unique_id, year_of_birth)][N>1, .(unique_id)]  
# duplicated_entries <- cust_by_yr[, .N, by = .(unique_id, year_of_birth, month_of_birth)][N>1, .(unique_id)]  
# duplicated_entries <- cust_by_yr[, .N, by = .(unique_id, year_of_birth, month_of_birth, gender_code)][N>1, .(unique_id)]  
# duplicated_entries <- cust_by_yr[, .N, by = .(unique_id, year_of_birth, month_of_birth, gender_code, ethnicity_code)][N>1, .(unique_id)]  
duplicated_entries <- cust_by_yr[, .N, by = .(unique_id, year_of_birth, month_of_birth, gender_code, ethnicity_code, customer_race_is_black_indicator)][N>1, .(unique_id)]  

pct_unduplicated_entries <- (nrow(cust_by_yr) - nrow(duplicated_entries)) / nrow(cust_by_yr)
pct_unduplicated_entries
```











