---
title: "OCS"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(magrittr)
R.utils::sourceDirectory("~/git/vecf_counts/src/functions/")
source(file = "~/git/vecf_counts/src/functions/fix_column_names.R")
source(file = "~/git/vecf_counts/src/functions/is_blank.R")

cust_by_yr <- fix_column_names(data.table::fread("~/git/vecf_counts/data/original/q4/OCS/OCS Services By Year.csv", colClasses = "character"))
values <- fix_column_names(data.table::fread("~/git/vecf_counts/data/original/ExportedValues.csv"))
```

####Notes: 

* I skipped [match_type] and [associated_id].  
+ match_type - missing in metadata, 100% of records have value "D"
+ associated_id - assuming this is an internal id?, 100% populated
* program year only 2016 
* service duration - this looks like a number reflecting days but potentially unclear and no entry in dictionary

```{r columns, echo=FALSE}
#. unique_id ----
pct_complete_unique_id <- (nrow(cust_by_yr) - cust_by_yr[is_blank(unique_id), .N]) / nrow(cust_by_yr)
#.. unique value count ----
unq_count_unique_id <- nrow(unique(cust_by_yr[, .(unique_id)]))
#.. valid values ----
wrong_id_length <- cust_by_yr[nchar(unique_id) != 7, .N]
pct_valid_values_unique_id <- (nrow(cust_by_yr) - wrong_id_length) / nrow(cust_by_yr)

#. indicator columns - clinical medication, dsmiv, medicaid ----
pct_complete_clin_med <- (nrow(cust_by_yr) - cust_by_yr[is_blank(clinical_medication), .N]) / nrow(cust_by_yr)
pct_complete_dsmiv <- (nrow(cust_by_yr) - cust_by_yr[is_blank(dsmiv), .N]) / nrow(cust_by_yr)
pct_complete_medicaid <- (nrow(cust_by_yr) - cust_by_yr[is_blank(medicaid_enrolled), .N]) / nrow(cust_by_yr)
#.. unique value count ----
unq_count_clin_med <- nrow(unique(cust_by_yr[, .(clinical_medication)]))
unq_count_dsmiv <- nrow(unique(cust_by_yr[, .(dsmiv)]))
unq_count_medicaid <- nrow(unique(cust_by_yr[, .(medicaid_enrolled)]))
#.. valid values ----
pct_valid_values_clin_med <- (nrow(cust_by_yr) - cust_by_yr[!clinical_medication %in% c("1", "2"), .N]) / nrow(cust_by_yr)
pct_valid_values_dsmiv <- (nrow(cust_by_yr) - cust_by_yr[!dsmiv %in% c("1", "2"), .N]) / nrow(cust_by_yr)
pct_valid_values_medicaid <- (nrow(cust_by_yr) - cust_by_yr[!medicaid_enrolled %in% c("1", "2"), .N]) / nrow(cust_by_yr)

#. date columns - program year, servoce begin date, service end date ----
pct_complete_prog_yr <- (nrow(cust_by_yr) - cust_by_yr[is_blank(program_year), .N]) / nrow(cust_by_yr)
pct_complete_serv_begin <- (nrow(cust_by_yr) - cust_by_yr[is_blank(service_begin_date), .N]) / nrow(cust_by_yr)
pct_complete_serv_end <- (nrow(cust_by_yr) - cust_by_yr[is_blank(service_end_date), .N]) / nrow(cust_by_yr)
#.. unique value count ----
unq_count_prog_yr <- nrow(unique(cust_by_yr[, .(program_year)]))
unq_count_serv_begin <- nrow(unique(cust_by_yr[, .(service_begin_date)]))
unq_count_serv_end <- nrow(unique(cust_by_yr[, .(service_end_date)]))
#.. valid values ----
pct_valid_values_prog_yr <- (nrow(cust_by_yr) - cust_by_yr[as.integer(program_year) < 2013, .N]) / nrow(cust_by_yr)
pct_valid_values_serv_beg <- (nrow(cust_by_yr) - 
                               cust_by_yr[stringr::str_sub(cust_by_yr$service_begin_date, start = 7L, end = 10L) < 2013, .N]) / 
                                nrow(cust_by_yr)
pct_valid_values_serv_end <- (nrow(cust_by_yr) - 
                               cust_by_yr[stringr::str_sub(cust_by_yr$service_end_date, start = 7L, end = 10L) < 2013, .N]) / 
                              nrow(cust_by_yr)

#. service columns - service duration, service placement ----
pct_complete_serv_dur <- (nrow(cust_by_yr) - cust_by_yr[is_blank(service_duration), .N]) / nrow(cust_by_yr)
pct_complete_serv_place <- (nrow(cust_by_yr) - cust_by_yr[is_blank(service_placement_type), .N]) / nrow(cust_by_yr)
#.. unique value count ----
unq_count_serv_dur <- nrow(unique(cust_by_yr[, .(service_duration)]))
unq_count_serv_place <- nrow(unique(cust_by_yr[, .(service_placement_type)]))
#.. valid values ----
pct_valid_values_serv_dur <- (nrow(cust_by_yr) - cust_by_yr[as.numeric(service_duration) < 1, .N]) / nrow(cust_by_yr)
pct_valid_values_serv_place <- (nrow(cust_by_yr) - cust_by_yr[is_blank(service_placement_type)|as.numeric(service_placement_type) <1, .N]) / nrow(cust_by_yr)


```



```{r UniqueID, echo = FALSE}
SummaryEval <- 
  rbind(
  tibble::tibble("Column" = "ID", "Measure" = c("% Complete", "# Unique", "% Valid "), 
                         "Value" = c(pct_complete_unique_id, 
                                     unq_count_unique_id, pct_valid_values_unique_id)),
    tibble::tibble("Column" = "Medication", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_clin_med, 
                               unq_count_clin_med, pct_valid_values_clin_med)),
    tibble::tibble("Column" = "DSMIV", "Measure"=c("% Complete", "# Unique", "% Valid "),
               "Value" = c(pct_complete_dsmiv, 
                           unq_count_dsmiv, pct_valid_values_dsmiv)),
    tibble::tibble("Column" = "Medicaid", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_medicaid, 
                               unq_count_medicaid, pct_valid_values_medicaid)),
    tibble::tibble("Column" = "Program Year", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_prog_yr, unq_count_prog_yr, 
                               pct_valid_values_prog_yr)),
    tibble::tibble("Column" = "Service Begin", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_serv_begin, unq_count_serv_begin, 
                               pct_valid_values_serv_beg)),
    tibble::tibble("Column" = "Service End", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_serv_end, unq_count_serv_end, 
                               pct_valid_values_serv_end)),
    tibble::tibble("Column" = "Duration", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_serv_dur, unq_count_serv_dur, 
                               pct_valid_values_serv_dur)),
    tibble::tibble("Column" = "Placement", "Measure"=c("% Complete", "# Unique", "% Valid "),
                   "Value" = c(pct_complete_serv_place, unq_count_serv_place, 
                               pct_valid_values_serv_place))
  ) #rbind

```

### Summary Table

```{R, echo = FALSE}
knitr::kable(reshape2::recast(SummaryEval, Column~Measure))
```

####Validations: 
* unique_id - filtered for nchar = 7
* clinical_medication - filtered for values 1 and 2
* dsmiv  - filtered for values 1 and 2
* medicaid_enrolled  - filtered for values 1 and 2
* program_year - filtered for years after 2013 (but not meaningful - all records are 2016)
* service_begin_date - filtered for years after 2013 (but not meaningful - all records are 2015-2016)
* service_duration - filtered out duration less than 1
* service_end_date - filtered for years after 2013 (but not meaningful - all records are 2015-2016)
* service_placement_type - filtered out blanks and values less than 1 



####Longitudinal checks

Total number of records is:

```{r, echo = FALSE}
print(nrow(cust_by_yr))
```

Since there are 9598 unique IDs - need to check persistence of traits for the duplicates. There don't seem to be demographic columns to check. All other columns could presumably change? Not sure how to proceed. There are 4241 duplicates. 

The percent unduplication rate is: 
```{r, echo = FALSE}
duplicated_entries <- cust_by_yr[, .N, by = .(unique_id)][N>1, .(unique_id)] 
# 4241 duplicates
# no demographic columns to check whether same person or not
pct_unduplicated_entries <- (nrow(cust_by_yr) - nrow(duplicated_entries)) / nrow(cust_by_yr)
pct_unduplicated_entries
```





